Diffusive Homeostasis in a Self-Organizing Recurrent Neural Network \\ \begin{large} Spatially dependent Interaction as a Determinant of Neural Activity and Plasticity \end{large}

### 2017-02-03 draft.tex
.

A constant "rate" of an approximately randomly spiking neuron means that the dependence of fluctuations of rates onto the width of the averaging window should be identical to a homogeneous poisson process of the same total mean firing rate. As bin sizes increase, possible long-term fluctuations of firing rates can then be identified as deviations from the variance one would expect from a homogeneous process. We tested this in Figure \ref{f_Var_vs_n_Bin}. (A) was taken from $\mathrm{1000-1500\,s}$, which is well inside the phase of diffusive homeostasis. As one can see, the statistics of the three representative neurons are well fitted by the predictions for homogeneous poisson processes with the same mean rate, being a straight line given by $\mathrm{var = f_{mean}\cdot n_{bins}/T_{total}}$, though slight deviations can be seen as a different factor of proportionality. (B) acts as an illustration showing a counterexample: Here, we included the entire time span of the simulation, including the transition between non-diffusive and diffusive homeostasis. For almost all of the neurons, this will cause a transition of firing rate between 3 Hz and a new rate different from 3 Hz. This inhomogeneity is represented by the increasing positive relative deviation from the dashed line for small number of bins (i.e., large bin widths), even increasing again for a very few number of bins.
\begin{figure}
%%% plot_spt_var_vs_bin_width.py
\includegraphics[width=\textwidth]{../../plots/Spike_Stats/sim_f_var_vs_hom_poisson.png}
\caption{A: Relation between Variance of the frequency signal (acquired by binning spikes) and the number of bins within the given time range of $\mathrm{1000-1500\,s}$, taken from three randomly picked excitatory neurons. B: Same procedure as in (A) for a randomly chosen excitatory neuron, but time range was $\mathrm{1000-1500\,s}$, which includes the initial phase of non-diffusive homeostasis. In both plots, dashed lines are the expected curve of a homogeneous poisson process of the same mean rate.}
\label{f_Var_vs_n_Bin}
\end{figure}
We therefore concluded that diffusive homeostasis manages to retain firing rates on a constant level, not only population-wise but on a single-cell level as well. In particular, plastic modifications across the network due to STDP and synaptic growth/pruning - which were always acting for all the simulation results presented so far - appeared to be compensated for.

@@

After having worked out the analytical basis, we compared the prediction obtained from numerically solving \eqref{NO_equil_cond_interact_matrix_operator} for certain spatial configurations of neurons to the steady-state firing rates of the full spiking network with the same spatial structure. In particular, we were interested in the quality of the predictions with respect to the choice of diffusion constant. Figure \ref{Rand_Matrix_Sol_vs_Sim} shows three examples: For $\mathrm{D=10\, \mu m^2 /s}$, the correlation between measured and predicted firing rates is very good. In contrast, we included the - obviously unsuccessful - attempt to predict firing rates for a simulation with instant diffusion based on the spatial structure by setting D to a relatively high value of $\mathrm{D=100\, \mu m^2 /s}$. Since instant diffusion overrides any spatial inhomogeneities, this represented a limiting case where our analytic model could not make any meaningful predictions. The outcome of the third case shown in the plot, $\mathrm{D=0}$, is correctly predicted by the analytic model. In general, instant diffusion as well as $\mathrm{D=0}$ overrode the effect of spatial heterogeneity onto firing rates. 

>>>>

A constant "rate" of an approximately randomly spiking neuron means that the dependence of fluctuations of rates onto the width of the averaging window should be identical to a homogeneous poisson process of the same total mean firing rate. As bin sizes increase, possible long-term fluctuations of firing rates can then be identified as deviations from the variance one would expect from a homogeneous process. We tested this in Figure \ref{f_Var_vs_n_Bin}. (A) was taken from $\mathrm{1000-1500\,s}$, which is well inside the phase of diffusive homeostasis. As one can see, the statistics of the three representative neurons are well fitted by the predictions for homogeneous poisson processes with the same mean rate, being a straight line given by $\mathrm{var = f_{mean}\cdot n_{bins}/T_{total}}$, though slight deviations can be seen as a different factor of proportionality. (B) acts as an illustration showing a counterexample: Here, we included the entire time span of the simulation, including the transition between non-diffusive and diffusive homeostasis. For almost all of the neurons, this will cause a transition of firing rate between 3 Hz and a new rate different from 3 Hz. This inhomogeneity is represented by the increasing positive relative deviation from the dashed line for small number of bins (i.e., large bin widths), even increasing again for a very few number of bins.
\begin{figure}
%%% plot_spt_var_vs_bin_width.py
\includegraphics[width=\textwidth]{../../plots/Spike_Stats/sim_f_var_vs_hom_poisson.png}
\caption{A: Relation between Variance of the frequency signal (acquired by binning spikes) and the number of bins within the given time range of $\mathrm{1000-1500\,s}$, taken from three randomly picked excitatory neurons. B: Same procedure as in (A) for a randomly chosen excitatory neuron, but time range was $\mathrm{0-1500\,s}$, which includes the initial phase of non-diffusive homeostasis. In both plots, dashed lines are the expected curve of a homogeneous poisson process of the same mean rate.}
\label{f_Var_vs_n_Bin}
\end{figure}
We therefore concluded that diffusive homeostasis manages to retain firing rates on a constant level, not only population-wise but on a single-cell level as well. In particular, plastic modifications across the network due to STDP and synaptic growth/pruning - which were always acting for all the simulation results presented so far - appeared to be compensated for.

@@

After having worked out the analytical basis, we compared the prediction obtained from numerically solving \eqref{NO_equil_cond_interact_matrix_operator} for certain spatial configurations of neurons to the steady-state firing rates of the full spiking network with the same spatial structure. In particular, we were interested in the quality of the predictions with respect to the choice of diffusion constant. Figure \ref{Rand_Matrix_Sol_vs_Sim} shows three examples: For $\mathrm{D=10\, \mu m^2 /ms}$, the correlation between measured and predicted firing rates is very good. In contrast, we included the - obviously unsuccessful - attempt to predict firing rates for a simulation with instant diffusion based on the spatial structure by setting D to a relatively high value of $\mathrm{D=100\, \mu m^2 /ms}$. Since instant diffusion overrides any spatial inhomogeneities, this represented a limiting case where our analytic model could not make any meaningful predictions. The outcome of the third case shown in the plot, $\mathrm{D=0}$, is correctly predicted by the analytic model. In general, instant diffusion as well as $\mathrm{D=0}$ overrode the effect of spatial heterogeneity onto firing rates. 
